### 2021-04-29 Paper/Main.tex
edit layout

\documentclass[12pt,oneside,reqno]{amsart}
\input{Paper/settings}

\begin{document}

%title page
\input{Paper/title_page}

\section{Introduction and Motivation}
\input{Paper/motivation}

\section{Literature Review}
\label{Literature}
\input{Paper/lit_review}

\section{Data}
\label{Data}
\input{Paper/data}

\section{Methodology}
\label{Methodology}
\input{Paper/Methodology}

\section{Results}
\label{Results}
\input{Paper/Results}

\section{Robustness Checks}
\input{Paper/robustness}
\label{Robustness}

\section{Conclusion}
\label{Conclusion}

\bibliography{Matt_Bib.bib}{}
\bibliographystyle{plain}
\singlespacing
\newgeometry{margin=.15in}

\section{Appendix}
%take these when needed
\begin{center}
    \paperspacingnarrow
    \input{Code/data/desc1}\\
    \textbf{Table 1: EXPLAIN THE TABLE!!!}
    \paperspacingwide
\end{center}

\begin{center}
    \paperspacingnarrow 
    \input{Code/data/size_summary_stat_size}\\
    \textbf{Table 2:} This table shows the mean results by decile before any regressions\\
    \paperspacingwide
 \end{center}
\vspace{1cm}

\begin{center}
    \paperspacingnarrow
    \input{Code/data/beta_output_inital}\\
    \textbf{Table 3:} Beta decile by estimated coefficients for the  five fama french factors and EMP factor\\
    \paperspacingwide
\end{center}

\vspace{1cm}

\begin{center}
    \paperspacingnarrow
    \input{Code/data/size_output_inital}\\
    \textbf{Table 4:} Size decile by estimated coefficients for the  five fama french factors and EMP factor\\
    \paperspacingwide
\end{center}

\begin{center}
    \paperspacingnarrow
    \input{Code/data/FF_pass_2}\\
    \textbf{Table 4:} second pass results for the size and beta decile portfolios\\
    \paperspacingwide
\end{center}

\begin{center}
    \paperspacingnarrow
    \input{Code/data/FF_pass_2_no_ESG}\\
    \textbf{Table 5:} second pass results for the size and beta decile portfolios without ESG factor\\
    \paperspacingwide
\end{center}

\begin{center}
    \paperspacingnarrow
    \input{Code/data/Industry_FF_pass_2}\\
    \textbf{Table 6:} second pass results for the size and beta decile portfolios without ESG factor\\
    \paperspacingwide

\end{center}

\clearpage
\restoregeometry

%\nocite{*}
%\printbibliography

%\section{Code}
%\input{Paper/Code_files}
\end{document}