### 2021-04-29 Paper/Methodology.tex

As discussed in the literature review, there are numerous models for assessing asset pricing and abnormal returns. In this paper, I will be using an extension of Fama French's 2015 paper ``A five-factor asset pricing model". This analysis method is built on the Fama French three factor model and relies on the two-pass sort with an assigned series of beta decile and size decile-based portfolios. Breaking the portfolios into these distinct groups allows for further analysis of the effect across these two factors to assess relational interdependency.

For this analysis, I sort each security into decile groups for beta and size (based on market cap). As firms' size and volatility are not constant functions, each firm is assigned a decile for each month depending on the most recent decile split.  Each security will be assigned a decile for beta and size according to the CRSP /NYSE American beta decile for beta and the CRSP Nasdaq Market Capitalization portfolio for size. After sorting each of the securities by beta and size decile, 300,264 of the original 830,889 data points had matches for both beta and size portfolios. To estimate the factor sensitivities base on the described decile groups, I ran the regression in equation \eqref{six_factor_basic_reg} on the cleaned data set.

\begin{equation}
\begin{split}
    R_{i,t}^e = a_i+b_i(R_{M,t}-R_{fr,t})+s_iSMB_t+ h_iHML_t+  r_iRMW_t+c_iCMA_t+e_iEMP_t +\epsilon_{i,t}
    \end{split}
    \label{six_factor_basic_reg}
\end{equation}

\begin{center}
Where:\\
    $R_{i,t}^e = R_{it}-R_{rf,t}$
\end{center}

Next, I used the data from equation \eqref{six_factor_basic_reg} to run the second regression in the two-pass sort. The second pass of the two-pass sort is described in the equation  \eqref{second_pass} and is used to obtain the risk premia. To consolidate the risk premium across time, a simple average as described in equation \eqref{summing_lamda} is used to collapse the risk premia for each of the given factors.

\begin{equation}
    E[R^e]= b\lambda_1+s\lambda_2 +h\lambda_3 +r\lambda_4 +c \lambda_4 + e\lambda_5
    \label{second_pass}
\end{equation}