### 2021-05-02 Paper/robustness.tex
latex update

when doing the SIC porlfios did not have data for SIC starting with 5 b/c they were all BB companies 
 %as dicussed in section \ref{Data} [talk about using mining instead of petrolum ]

 dicuss limitation of only one set of ESG factors for MSCI

 we used the Size deciles on a go froward basis when analysis the data as the statistical relationship was more apparent

 \textbf{Analysis for the w/o ESG on the data}\\
 \textbf{yearly effect}\\

 \textbf{MSCI data 2019 beyond}\\

 \textbf{MSCI SIC level analysis}\\

 problems we can never solve for both of these indexes does not do a great job capturing small companies so size might be a bit of a probelm overall as the small companies do not have ther ESG scores baked in as MSCI does not have the resources to rank every company