### 2021-05-02 Paper/Main.tex
Updates from Overleaf

\documentclass[12pt,oneside,reqno]{amsart}
\input{Paper/settings}

\begin{document}

%title page
\input{Paper/title_page}

\section{Introduction and Motivation}
\input{Paper/motivation}

\section{Literature Review}
\label{Literature}
\input{Paper/lit_review}

\section{Data}
\label{Data}
\input{Paper/data}

\section{Methodology}
\label{Methodology}
\input{Paper/Methodology}

\section{Results}
\label{Results}
\input{Paper/Results}

\section{Robustness Checks}
\input{Paper/robustness}
\label{Robustness}

\section{Conclusion}
\label{Conclusion}

\bibliography{Matt_Bib.bib}{}
\bibliographystyle{plain}
\singlespacing
\newgeometry{margin=.15in}

\section{Appendix}
%take these when needed

\begin{center}
    \paperspacingnarrow
    \input{Code/data/FF_pass_2_no_ESG}\\
    \textbf{Table 5:} second pass results for the size and beta decile portfolios without ESG factor\\
    \paperspacingwide
\end{center}

\begin{center}
    \paperspacingnarrow
    \input{Code/data/Industry_FF_pass_2}\\
    \textbf{Table 6:} second pass results for the size and beta decile portfolios without ESG factor\\
    \paperspacingwide
\end{center}

\begin{center}
    \paperspacingnarrow
    \input{Code/data/Time_FF_pass_2}\\
    \textbf{Table 7:} Time based cofficents
    \paperspacingwide
\end{center}

\begin{center}
    \paperspacingnarrow
\input{Code/data/All_Industry_FF_pass_2}\\
\textbf{Table 7:} SIC top 10\% minus bottom 10\%
    \paperspacingwide
\end{center}

\begin{center}
    \paperspacingnarrow
\input{Code/data/All_Industry_FF_pass_2_past_2018}\\
\textbf{Table XX:} SIC top 10\% minus bottom 10\% 2018-2020
    \paperspacingwide
\end{center}
\clearpage
\restoregeometry

%\nocite{*}
%\printbibliography

%\section{Code}
%\input{Paper/Code_files}
\end{document}