### 2021-05-02 Paper/Results.tex

Table 2 represents a matrix of values for the first pass in the Fama French two-pass sort, described in equation \eqref{six_factor_basic_reg}. The row axis is the distinct size deciles, where 1 is the smallest 10\% of companies by market cap companies, and 10 is the decile with the largest 10\% of companies by market cap. The column axis includes estimated coefficients for the intercept, five Fama French factors, and the EMP factor. $HML_{t}$ is the difference between the returns on diversified portfolios of the high and low book to market stocks.  $SMB_{t}$ is the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks. $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability. $CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms. $EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index. Finally, the intercept is the excess returns for the given decile. The results in this table demonstrate that for HML, CMA, and EMP factors, there is a size-based relationship wherein the larger the size decile, the smaller the sensitivity effect is on the factor. In addition, the EMP factor also exhibits a consistent negative relationship between the EGS factor and returns. This relationship indicates that investors are returning a premium by investing in ESG companies over the contra ESG counterparts and that this relationship is largest for small companies. The results from this statistical test are significant at the 5\% level for the EMP factor for all of the portfolios except the smallest and largest size portfolios. However, several of the Fama French factors are also not significant for the largest and smallest portfolios, which may indicate that these tail portfolios might be behaving differently due to their composition rather than a reflection of the factor.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/size_output_inital}\\
    \textbf{Table 2:} represents the first pass of the two pass sort where the portfolios are broken up by size. the row axis is portfolio and the column axis is the  market beta,four Fama French factors and the EMP factor\\
    \paperspacingwide
\end{center}

Table 3 demonstrates the risk premia for each of the factors using the second step in the two-pass sort for Size decile portfolios. The market factor, the SMB factor, the CMA factor, and the EMP factor are all positive and statistically significant at the 1\% level. The other two factors have a negative sign and are not statistically significant at the 10\% level.  EMP factor demonstrates that a 1 unit change in beta corresponds with an 8.7\% increase in returns. Although this is by far the smallest coefficient of the four statistically significant returns, this term is still economically meaningful and represents a risk premia for investing in high ESG firms. Overall the model appears to be a fairly good fit with a $R^2$ value of 87.5\%, which is lower than some previous literature using the petroleum index but still indicates a well fit model \hl{cite this paper}

\begin{center}
    \paperspacingnarrow
    \input{Code/data/FF_pass_2}\\
    \textbf{Table 3:} second pass results for the size and beta decile portfolios\\
    \paperspacingwide
\end{center}

Table 4 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the row axis is the beta decile. In this table,  the beta decile 1 comprises the companies with the lowest market beta's. Similarly, beta decile 10 is the portfolio of companies with the largest market beta values.  Unlike the size decile portfolios, the beta decile portfolios do not have consistently negative relations with respect to decile. Rather the relationship appears to have a dome shape where portfolio 6 has the smallest positive sensitivity to the EMP factor, and larger or smaller beta portfolios appear to have a larger sensitivity to the  EMP factor. \hl{why might this be}. Similar to the size decile analysis, the EMP coefficients are all very significant, with every portfolio having significance at the 5\% level.

When considering the second phase of the two-pass sort for the beta decile portfolios, the similarities to the size decile continue wherein market returns, SMB, CMA, and EMP exhibit positive risk premia. The other two factors remain statistically insignificant, and the MHL factor changed signs and now has a positive relationship. The EMP coefficient indicates that a one-unit change in beta would result in a 7.36\% change in returns for a high ESG company. This result is slightly smaller than the result from the size portfolio but still demonstrates a statistical relationship in the same order of magnitude. The model is also a good fit with an $R^2$ value of 86.0\%. With the baseline analysis proving that there is a relationship between ESG and excess marker returns, I will provide an investigation of the robustness in section \ref{Robustness} by first comparing the results of the two-pass sort with the EGS factor to the same analysis without the factor. Next, I will investigate time-related relationships between ESG and excess returns by implementing a dummy variable for the year and repeating the two-pass sort methodology.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/beta_output_inital}\\
    \textbf{Table 4:} Beta decile by estimated coefficients for the  five fama french factors and EMP factor\\
    \paperspacingwide
\end{center}

Moving on to the second stage of the analysis, I constructed an ESG factor out of the MSCI ESG ratings to capture a more holistic view of ESG risk. I created this factor by taking the top 10\% of MSCI EGS rated companies and subtracted them from the bottom 10\% of firms rated by the MSCI. I repeated the methodology above for the regression analysis. I  regressed two sets of two-pass regression analyses, one on beta-sized portfolios and another on size-based portfolios. In the first pass results for the MSCI factor, the beta portfolio demonstrated a decrease in sensitivity as the betas grew; however, the size-based portfolios did not experience any real pattern governing the decile and the sensitivity. In the case of the size-based portfolios, some of the sensitivity figures were positive while others were negative. Table 5 represents a second stage two-pass sort regression output. Similar to the regression in the first stage of the analysis, the market return, CMA, SMB remain statistically significant; however, only the size-based portfolios demonstrate a statistically significant coefficient for the ESG factor.  The EMP ESG factor coefficient can be interpreted as a negative relationship between ESG and excess return wherein a one beta unit increase in ESG results in a decrease of 2\% in excess returns. One limitation of the MSCI data is it only represents a snapshot in ESG rating data; that is, the MSCI data is only the most recent ESG performance and does not account for changes in ESG performance. As such, a company that used to be a  poor ESG and has improved to a strong ESG company will only be categorized as a strong ESG company which will not accurately address the ESG performance.  As such, in section \ref{Robustness} I will address this by shrinking the time frame to see if the coefficients increase in their significance. Further,  to address additional concerns of robustness, I will run another two-pass sort to isolate the SIC industry-specific ESG factor and determine if there are ESG related excess returns that are greater in some industries than others.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/All_Industry_FF_pass_2}\\
    \textbf{Table 5:} SIC top 10\% minus bottom 10\%
    \paperspacingwide
\end{center}