### 2021-05-03 Paper/Main.tex
latex now basically to the end!!

\documentclass[12pt,oneside,reqno]{amsart}
\input{Paper/settings}

\begin{document}

%title page
\input{Paper/title_page}

\section{Introduction and Motivation}
\input{Paper/motivation}

\section{Literature Review}
\label{Literature}
\input{Paper/lit_review}

\section{Data}
\label{Data}
\input{Paper/data}

\section{Methodology}
\label{Methodology}
\input{Paper/Methodology}

\section{Results}
\label{Results}
\input{Paper/Results}

\section{Robustness Checks}
\input{Paper/robustness}
\label{Robustness}

\section{Conclusion}
\label{Conclusion}
\input{Paper/conclusion}

\bibliography{Matt_Bib.bib}{}
\bibliographystyle{plain}

%\nocite{*}
%\printbibliography

%\section{Code}
%\input{Paper/Code_files}
\end{document}