### 2021-05-03 Paper/robustness.tex

when doing the SIC porlfios did not have data for SIC starting with 5 b/c they were all BB companies 
 %as dicussed in section \ref{Data} [talk about using mining instead of petrolum ]

 \begin{center}
    \paperspacingnarrow
    \input{Code/data/FF_pass_2_no_ESG}\\
    \textbf{Table 6:} second pass results for the size and beta decile portfolios without ESG factor\\
    \paperspacingwide
\end{center}

 dicuss limitation of only one set of ESG factors for MSCI

 we used the Size deciles on a go froward basis when analysis the data as the statistical relationship was more apparent

 \textbf{Analysis for the w/o ESG on the data}\\
 \textbf{yearly effect}\\
\begin{center}
    \paperspacingnarrow
    \input{Code/data/Time_FF_pass_2}\\
    \textbf{Table 7:} Time based cofficents
    \paperspacingwide
\end{center}
 \textbf{MSCI data 2018-2020}\\
 \begin{center}
    \paperspacingnarrow
\input{Code/data/All_Industry_FF_pass_2_past_2018}\\
\textbf{Table XX:} SIC top 10\% minus bottom 10\% 2018-2020
    \paperspacingwide
\end{center}
 \textbf{MSCI SIC level analysis}\\

  \begin{center}
    \paperspacingnarrow
    \input{Code/data/Industry_FF_pass_2}\\
    \textbf{Table 6:} second pass results for the size and beta decile portfolios without ESG factor\\
    \paperspacingwide
\end{center}

 WE ADJUSTED THE YEAR to 2019!!
 problems we can never solve for both of these indexes does not do a great job capturing small companies so size might be a bit of a probelm overall as the small companies do not have ther ESG scores baked in as MSCI does not have the resources to rank every company