### 2021-05-03 Paper/robustness.tex

As discussed in section \ref{Results}, additional tests are required to ensure the methodology's robustness. The first robustness check I will run is a repetition of the two-pass sort for where the ESG factor is calculated asMSCI ESG index minus the petroleum index (EMP). Table 6 confirms the statistical significance of the market term, the SMB term, and the CMA term. All three of these terms were found to be statistically significant in the initial regression as well. Beyond the coefficients,  the results in table 6 demonstrate a slightly lower $R^2$ for both the Beta decile and the size decile variants, demonstrating that the EMP term contributes to the explanation of excess returns. In addition, the constant term for both size and beta decile portfolios increased in magnitude as a result of omitting the EMP. As the coefficient term gets closer to zero, the factors do a better job describing the returns and indicate that there is no excess return that is not explained. 
 %as dicussed in section \ref{Data} [talk about using mining instead of petrolum ]

 \begin{center}
    \paperspacingnarrow
    \input{Code/data/FF_pass_2_no_ESG}\\
    \textbf{Table 6:} second pass results for the size and beta decile portfolios without ESG factor\\
    \paperspacingwide
\end{center}
Another robustness check that is essential to consider is the time-based effects on the model. Although we have demonstrated that the EMP factor is statistically significant and explains part of our model, it is essential to understand if this is ubiquitous across time in the sample or if the  ESG factor has shifted in risk premia over time. To address this robustness concern, I have repeated the two-pass sort with the size decile portfolio by adding a time-based dummy variable. Table 7 details the results from the regressions. In 2015 2016, 2018, and 2019 there is statistical significance for the EMP term and a positive association between EMP and excess returns. The ESG factor showed the most strength in 2019 as well as the greatest coefficient value. In 2019 a one-unit change in beta would result in a 25.9\% increase in excess returns. The coefficient values in 2017 appeared to hold a negative relationship as a one-unit change in ESG beta would have resulted in a reduction in excess returns by 10.5\%. Although in 2020, there was a positive relationship, the coefficient is not statistically significant. However, in 2020, the CMA factor, which has been statistically significant at the 1\% level for all other tests, also does not appear to be statistically significant. 2020 represents a bit of a deviation due to the larger market condition from the COVID-19 pandemic rather than the EMP factor's long-run behavior.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/Time_FF_pass_2}\\
    \textbf{Table 7:} Shows the results for the second pass in a two pass sort. The two pass sort used size based deciles for the first pass. Columns (1)-(6) represent individual years of data. 
    \paperspacingwide
\end{center}
 \textbf{MSCI data 2018-2020}\\
 \begin{center}
    \paperspacingnarrow
\input{Code/data/All_Industry_FF_pass_2_past_2018}\\
\textbf{Table 8:} SIC top 10\% minus bottom 10\% 2018-2020
    \paperspacingwide
\end{center}
 \textbf{MSCI SIC level analysis}\\

  \begin{center}
    \paperspacingnarrow
    \input{Code/data/Industry_FF_pass_2}\\
    \textbf{Table 6:} second pass results for the size and beta decile portfolios without ESG factor\\
    \paperspacingwide
\end{center}

 WE ADJUSTED THE YEAR to 2019!!
 problems we can never solve for both of these indexes does not do a great job capturing small companies so size might be a bit of a probelm overall as the small companies do not have ther ESG scores baked in as MSCI does not have the resources to rank every company

 when doing the SIC porlfios did not have data for SIC starting with 5 b/c they were all BB companies

dicuss limitation of only one set of ESG factors for MSCI

 we used the Size deciles on a go froward basis when analysis the data as the statistical relationship was more apparent