### 2021-05-05 Paper/title_page.tex

\begin{titlepage}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % Defines a new command for the horizontal lines, change thickness here

\center % Center everything on the page

%----------------------------------------------------------------------------------------
%	HEADING SECTIONS
%----------------------------------------------------------------------------------------

\textsc{\LARGE Colgate University}\\[1.5cm] % Name of your university/college
\includegraphics[scale=.1]{Paper/secondary/colgate_logo.png}\\[1cm] % Include a department/university logo - this will require the graphicx package
\textsc{\Large Financial Economics Seminar}\\[0.5cm] % Major heading such as course name
\textsc{\large Econ 487}\\[0.5cm] % Minor heading such as course title

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\HRule \\[0.4cm]
{ \huge \bfseries Factor Investing with ESG}\\[0.4cm] % Title of your document
\HRule \\[1.5cm]

%----------------------------------------------------------------------------------------
%	AUTHOR SECTION
%----------------------------------------------------------------------------------------

\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
Author: \emph{}
Matthew Chistolini \textsc{}\\ % Your name
\end{flushleft}

\end{minipage}\\[1cm]

% If you don't want a supervisor, uncomment the two lines below and remove the section above
 Advisor: \emph{}
Michael Connolly \textsc{}\\[2cm] % Your name

%----------------------------------------------------------------------------------------
%	DATE SECTION
%----------------------------------------------------------------------------------------

{\large \today}\\[2cm] % Date, change the \today to a set date if you want to be precise

\vfill % Fill the rest of the page with whitespace

\end{titlepage}