### 2021-05-07 Paper/Results.tex

Table 2 represents a matrix of values for the first pass in the Fama French two-pass sort, described in equation \eqref{six_factor_basic_reg}. The row axis is the distinct size deciles, where 1 is the smallest 10\% of companies by market cap companies, and 10 is the decile with the largest 10\% of companies by market cap. The column axis includes estimated coefficients for the five Fama French factors and the EMP factor. $HML_{t}$ is the difference between the returns on diversified portfolios of the high and low book to market stocks.  $SMB_{t}$ is the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks. $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability. $CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms. $EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index. The results in this table demonstrate that for HML, CMA, and EMP factors, there is a size-based relationship wherein the larger the size decile, the smaller the sensitivity effect is on the factor.  This relationship indicates that investors are returning a premium by investing in ESG companies over the contra ESG counterparts and that this relationship is strongest for small companies.  The size-based sensitivity that is greatest for small firms was also noted in Dorfleitner et al's work\cite{Dorfleitner2020ESGCA}. The results from this statistical test are significant at the 5\% level for the EMP factor for all of the portfolios except the smallest and largest size portfolios. However, several of the Fama French factors are also not significant for the largest and smallest portfolios. This finding indicate that these tail portfolios  behave differently and may require winsorizing rather than a reflection of the factor's significance.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/size_output_inital}\\
    \textbf{Table 2:} represents the first pass of the two pass sort where the portfolios are broken up by size deciles. The row axis is portfolio and the column axis is the  market beta, four Fama French factors, and the EMP factor.\\
    \paperspacingwide
\end{center}
Table 3 demonstrates the risk premia for each of the factors using the second step in the two-pass sort for size decile portfolios. The market factor, the SMB factor, the CMA factor, and the EMP factor are all positive and statistically significant at the 1\% level. The other two factors have a negative sign and are not statistically significant at the 10\% level.  EMP factor demonstrates that a 1 unit change in beta corresponds with an 8.7\% increase in excess returns. Although this is by far the smallest coefficient of the four statistically significant factors, this term is still economically meaningful and represents a risk premium can be achieved by investing in high ESG firms. Overall the model appears to fit the data reasonably well with a $R^2$ value of 87.5\%, which is lower than some previous literature using the petroleum index but still indicates a well-fit model \cite{Gregory2020TheGP}.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/FF_pass_2}\\
    \textbf{Table 3:} shows the results for the second pass of the two pass sort with the EMP factor. On the left labeled as (1) are the regression results for the size decile portfolios. On the right labeled as (2) are the regression results for the beta decile portfolios.\\
    \paperspacingwide
\end{center}

Table 4 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the row axis is the beta decile. In this table,  the beta decile 1 comprises the companies with the lowest market beta's. Similarly, beta decile 10 is the portfolio of companies with the largest market beta values.  Unlike the size decile portfolios, the beta decile portfolios do not have consistently negative relations with respect to decile. Rather the relationship appears to have a dome shape where portfolio 6 has the smallest positive sensitivity to the EMP factor, and larger or smaller beta portfolios appear to have a larger sensitivity to the  EMP factor. Similar to the size decile analysis, the EMP coefficients are all highly significant, with every portfolio having significance at the 5\% level.

When considering the second phase of the two-pass sort for the beta decile portfolios, the similarities to the size decile continue wherein market returns, SMB, CMA, and EMP exhibit positive risk premia and the other two factors remain statistically insignificant. The EMP coefficient indicates that a one-unit change in beta would result in a 7.36\% change in excess returns for a high ESG company. This result is slightly smaller than the result from the size portfolio but still demonstrates a statistical relationship in the same order of magnitude and suggests the EGS factor can provide excess returns. The beta decile model is a strong fit with an $R^2$ value of 86.0\%.

The two-pass sort indicates that there is a measurable relationship between the EMP factor and excess marker returns. In section \ref{Robustness} I will provide an analysis of the robustness by first comparing the results of the two-pass sort with the EGS factor to the same analysis without the factor. Next, I will investigate the time-related relationships between ESG and excess returns by implementing a dummy variable for the year and repeating the two-pass sort methodology.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/beta_output_inital}\\
    \textbf{Table 4:} shows the first pass in a two pass sort. The portfolios are decomposed by beta deciles and the table demonstrates the sensitivity of the four factors, market returns and and EMP factor to the decile.\\
    \paperspacingwide
\end{center}

Moving on to the second stage of the analysis, I constructed an ESG factor out of the MSCI ESG ratings to capture a more holistic view of ESG risk. I created this factor by taking the top 10\% of MSCI EGS rated companies and subtracted them from the bottom 10\% of firms rated by the MSCI. I repeated the methodology above for the regression analysis. I  regressed two sets of two-pass sort  regression analyses, one on beta-sized portfolios and another on size-based portfolios. In the first pass results for the MSCI ESG factor, the beta portfolio demonstrated a decrease in sensitivity as the beta deciles increased. However, the size-based portfolios did not experience any pattern, with some of the coefficients were positive while others were negative.

Table 5 represents the second stage in two-pass sort regression output. Similar to the regression with the EMP factor, the market return, CMA, SMB remain statistically significant. However, neither the size based size-based or beta-based portfolios demonstrate a statistically significant coefficient for the ESG factor. This lack of statistical significance indicates that the MSIC ESG factor is priced into one of the other statistically significant factors and should not produce any excess returns. Further solidifying that the factor is not economically significant is the high $R^2$  value for both of the portfolios, which indicates the strong fit of the model.

One potential reason that there is no statistical significance for the factor is the limitation of the MSCI data. MSCI only provides current  MSCI ESG rating data. As such, a company that used to be a  poor ESG firm that has improved to a strong ESG company will only be categorized as a strong ESG company which will not accurately address the ESG performance.

\begin{center}
    \paperspacingnarrow

    \input{Code/data/All_Industry_FF_pass_2}\\
    \textbf{Table 5:} shows the results for the second pass of the two-pass sort with the ESG factor generated from the top 10\% of MSCI ESG scores minus bottom 10\% MSCI ESG scores. On the left labeled as (1) are the regression results for the size decile portfolios. On the right, labeled as (2) are the regression results for the beta decile portfolios.

    \paperspacingwide
\end{center}

In section \ref{Robustness} I will address the data concerns by reducing the time frame and observing if the coefficients increase in significance. Further, to address additional concerns of robustness, I will run another two-pass sort on individual industries to see if there is uniformity to the risk prima response to the MSCI ESG factor.