### 2021-05-07 Paper/abstract.tex

\pagebreak
\hspace{0pt}
\vfill
\begin{center}
\Huge{Abstract}\\
\normalsize
\vspace{2cm}
ESG investing has more than one definition for  ``good" and ``bad" firms. This paper uses the Fama French five-factor model framework to explore two definitions of EGS risk premia and find radically different results.  When defining the ESG factor as the difference between the MSCI ESG index and the S\&P petroleum index, there is robust evidence for an economically and statistically significant result. This paper finds that a one-unit change in beta results in an 8.7\% increase in excess returns for size decile portfolios. However, changing the ESG factor to the difference between high ESG firms minus low ESG firms as rated by MSCI, there are no statistically measurable risk premia. This lack of uniformity underpins the difficulty of setting uniform policy, which can have a meaningful impact on millions of investor's retirement accounts.  
\end{center}
\vfill
\hspace{0pt}
\pagebreak